### 2017-06-15 analysis.Rmd
wmatrix images

---
title: "FocalGstar"
author: "Viliam Simko"
date: "June 2, 2017"
output:
  pdf_document: default
  html_document: default
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r setup, echo=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(rgdal)
library(readr)
library(corrplot)
library(plotrix)

knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "images/gen-"
)

source("R/hotspot_analysis_lib.R")
set.seed(0)

ZOOM_WMATRIX_SIZE <- 11
FOCAL_MATRIX_SIZE <- 41
MAX_AGGREG_LEVEL <- 8
ZOOMLEVEL <- 3
COLORMAP <- interpolate_colortable(
  howmany = 100, "black", "blue", "yellow", "red")

```

## Loading hi-res raster

Generating new rasters by aggregation
```{r loadraster, cache=TRUE}
hires <- raster("data/hires-raster.tif")

rast <- list()
rast[[1]] <- hires
for (i in 2:MAX_AGGREG_LEVEL) {
  rast[[i]] <- aggregate(hires, fact = i, fun = sum, na.rm = TRUE)
}
```

Rendering the aggregated rasters:
```{r rawdata, cache=TRUE, fig.width=5, fig.height=5}
op <- par(mar = rep(0, 4))
for (i in seq_along(rast)) {
  plot(log(rast[[i]]), xlab = NA, ylab = NA,
       legend = FALSE, axes = FALSE, box = FALSE,
       col = COLORMAP, colNA = "black")
}
par(op)
```
## Demo: Variable resolution (zoom)

### Weight matrix
```{r zoom-wmatrix, echo=FALSE, cache=TRUE, fig.width=4, fig.height=4}
ZOOM_WMATRIX <- weight_matrix_circular_fade(ZOOM_WMATRIX_SIZE, 0)
ZOOM_WMATRIX[is.na(ZOOM_WMATRIX)] <- 0 # removing any NAs from the weight matrix
plot_weight_matrix(ZOOM_WMATRIX)
```

### Standard G* (zoom)

First, we compute all the gstar rasters.
```{r, echo=FALSE, cache=TRUE}
zoom_gstar <- list()
pb <- txtProgressBar(max = length(rast), style = 3)
for (i in 1:length(rast)) {
  setTxtProgressBar(pb, i)
  r <- rast[[i]]
  zoom_gstar[[i]] <- GetisOrd(r, ZOOM_WMATRIX)
}
close(pb)
```
This code can be used for saving rasters as geotiffs, useful for external processing.
We don't execute the code when generating this document.
```{r, echo=FALSE, eval=FALSE}
z <- 0
for (g in zoom_gstar) {
  z <- z + 1
  writeRaster(g, sprintf("data/zoom/gstar-%03dx%03d-z%02d-w%d.tif",
                         nrow(g), ncol(g), z, nrow(ZOOM_WMATRIX)),
              overwrite = TRUE)

}
```

Now, we want to generate plain images out of the G* rasters:
```{r raw-zoom-gstar, echo=FALSE, cache=TRUE, fig.width=4, fig.height=4}
op <- par(mar = rep(0, 4))
for (g in zoom_gstar) {
  plot(g, xlab = NA, ylab = NA,
       legend = FALSE, axes = FALSE, box = FALSE, col = COLORMAP, colNA = "black")
}
par(op)
```

```{r demo-zoom-gstar, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4.5}
z <- 0
for (g in zoom_gstar) {
  z <- z + 1
  plot_zscore(g,
              title = sprintf("Standard G*: zoom=%d", z),
              sigma_thresh = 2)
}
```

### Focal G* (zoom)

First, we compute all the Focal G* rasters.
```{r, echo=FALSE, cache=TRUE}
zoom_focalgstar <- list()
pb <- txtProgressBar(max = length(rast), style = 3)
for (i in seq_along(rast)) {
  setTxtProgressBar(pb, i)
  r <- rast[[i]]
  st <- GetisOrdLocalStats(r, weight_matrix_circular(FOCAL_MATRIX_SIZE))
  zoom_focalgstar[[i]] <- GetisOrd(r, ZOOM_WMATRIX, st)
}
close(pb)
```

This code can be used for saving rasters as geotiffs, useful for external processing.
We don't execute the code when generating this document.
```{r, echo=FALSE, eval=FALSE}
z <- 0
for (g in zoom_focalgstar) {
  z <- z + 1
  writeRaster(g, sprintf("data/zoom/focalgstar-%03dx%03d-z%02d-w%d.tif",
                         nrow(g), ncol(g), z, nrow(ZOOM_WMATRIX)),
              overwrite = TRUE)

}
```

Now, we want to generate plain images out of the Focal G* rasters:
```{r raw-zoom-focalgstar, echo=FALSE, cache=TRUE, fig.width=4, fig.height=4}
op <- par(mar = rep(0, 4))
for (g in zoom_focalgstar) {
  plot(g, xlab = NA, ylab = NA, legend = FALSE, axes = FALSE, box = FALSE,
       col = COLORMAP, colNA = "black")
}
par(op)
```

```{r demo-zoom-focalgstar, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4.5}
z <- 0
for (g in zoom_focalgstar) {
  z <- z + 1
  plot_zscore(g,
              title = sprintf("Focal G*: zoom=%d", z),
              sigma_thresh = 2)
}
```

## Demo: Variable weight matrix (blur)
### Weight matrix sizes
```{r}
r <- rast[[ZOOMLEVEL]]
wsizes <- (1:12)*2 + 5
wsizes # printed

# selection to be rendered for latex inclusion
RENDERED_IDX <- floor(seq(1, length(wsizes), length.out = 8))
```

We render `r length(RENDERED_IDX)` weight matrices that will be included into a LaTeX document. In order to do this, we selected indices `r RENDERED_IDX`:
```{r blur-wmatrices, echo=FALSE, cache=TRUE, fig.width=4, fig.height=4}
for (ws in wsizes[RENDERED_IDX]) {
  w <- weight_matrix_circular_fade(ws, 0)
  w[is.na(w)] <- 0 # removing any NAs from the weight matrix
  plot_weight_matrix(w, show.minmax = FALSE)
}
```

### Standard G* (blur)
```{r demo-blur-gstar, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4.5}
for (ws in wsizes) {

  w <- weight_matrix_circular_fade(ws, 0)
  w[is.na(w)] <- 0 # removing any NAs from the weight matrix

  g <- GetisOrd(r, w)
  plot_zscore(g,
              title = sprintf("Standard G*: wsize=%d", ws),
              sigma_thresh = 2)
  writeRaster(g,
              sprintf("data/blur/gstar-%03dx%03d-z%02d-w%02d.tif",
                      nrow(r), ncol(r), ZOOMLEVEL, ws),
              overwrite = TRUE)
}
```

### Focal G* (blur)
```{r demo-blur-focalgstar, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4.5}
st <- GetisOrdLocalStats(r, weight_matrix_circular(FOCAL_MATRIX_SIZE))
for (ws in wsizes) {

  w <- weight_matrix_circular_fade(ws, 0)
  w[is.na(w)] <- 0 # removing any NAs from the weight matrix

  g <- GetisOrd(r, w, st)
  plot_zscore(g,
              title = sprintf("Focal G*: wsize=%d", ws),
              sigma_thresh = 2)
  writeRaster(g,
              sprintf("data/blur/focalgstar-%dx%d-z%02d-w%02d.tif",
                      nrow(r), ncol(r), ZOOMLEVEL, ws),
              overwrite = TRUE)
}
```

## Example of Parent-Child Clumping Difference
```{r metric-example, message=FALSE, echo=FALSE, cache=TRUE, fig.width=7, fig.height=4.5}
r <- rast[[ZOOMLEVEL]]
st1 <- GetisOrdStandardStats(r)
st2 <- GetisOrdLocalStats(r, weight_matrix_circular(FOCAL_MATRIX_SIZE))
plot_clumping_diff(r, st1, ZOOM_WMATRIX_SIZE, 29)
plot_clumping_diff(r, st2, ZOOM_WMATRIX_SIZE, 29)
```

```{r, echo=FALSE, message=FALSE}
glob <- read_csv("data/heatmap/aggregatedGlobal.csv")
foc <- read_csv("data/heatmap/aggregatedFocal.csv")
```

```{r, eval=FALSE, echo=FALSE}
foc %>%
  cor(use = "pairwise.complete.obs") %>%
  corrplot.mixed(order = "FPC", diag = "n", outline = "gray")
```

```{r, eval=FALSE, message=FALSE}
glob %>%
  select(W,Z,Down,Up) %>%
  cor(use = "pairwise.complete.obs") %>%
  corrplot.mixed(diag = "n", outline = "gray")
```

```{r zoom-sohup, echo=FALSE, eval=TRUE, cache=TRUE, message=FALSE}
zoom_soh <- read_csv("data/zoom_soh.csv")

zoom_soh %>% dplyr::select(Zoom, "G*" = UpZoom) -> y1
zoom_soh %>% dplyr::select(Zoom, "Focal G*" = UpFocalZoom) -> y2

plot(y1, type = "n", ylim = c(0,1),
     ylab = expression(SoH[up]),
     xlab = "Zoom level")

lines(y1, col = "blue", lw = 2)
lines(y2, col = "red", lw = 2)
legend(1,1, c("G*", "Focal G*"), fill = c("blue", "red"))
mtext("CI: FMsize and WMsize", line = -1, adj = .98)
```

```{r zoom-sohdown, echo=FALSE, eval=TRUE, cache=TRUE, message=FALSE}
zoom_soh <- read_csv("data/zoom_soh.csv")

zoom_soh %>% dplyr::select(Zoom, "G*" = DownZoom) -> y1
zoom_soh %>% dplyr::select(Zoom, "Focal G*" = DownFocalZoom) -> y2

plot(y1, type = "n", ylim = c(0,1),
     ylab = expression(SoH[down]),
     xlab = "Zoom level")

lines(y1, col = "blue", lw = 2)
lines(y2, col = "red", lw = 2)
legend(1,1, c("G*", "Focal G*"), fill = c("blue", "red"))
mtext("CI: FMsize and WMsize", line = -1, adj = .98)
```

```{r weight, echo=FALSE, eval=TRUE, cache=TRUE, message=FALSE}
d3 <- read.csv("data/w_focal_trued3.csv")
d3global <- read.csv("data/w_focal_falsed3.csv")

mean <- aggregate(d3$Up, by = list(d3$W + 4), mean)
sd <- aggregate(d3$Up, by = list(d3$W + 4), sd)
meanG <- aggregate(d3global$Up, by = list(d3global$W + 4), mean)
sdG <- aggregate(d3global$Up, by = list(d3global$W + 4), sd)

plot(NA,
     ylab = expression(SoH^Up),
     xlab = sprintf("Weight matrix size"),
     xlim = c(7,43), ylim = c(0,1))
plotCI(x = mean$Group.1,
       y = mean$x,
       uiw = sd$x, err = "y", pch = 20, slty = 3, scol = "red", add = TRUE)
plotCI(x = meanG$Group.1,
       y = meanG$x,
       uiw = sdG$x, err = "y", pch = 20, slty = 3, scol = "blue", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
lines(meanG$Group.1, meanG$x, col = "blue", lw = 4)
legend("topleft", c("Focal G*", "G*") , fill = c("red", "blue"))

mean <- aggregate(d3$Down, by = list(d3$W + 4), mean)
sd <- aggregate(d3$Down, by = list(d3$W + 4), sd)
meanG <- aggregate(d3global$Down, by = list(d3global$W + 4), mean)
sdG <- aggregate(d3global$Down, by = list(d3global$W + 4), sd)

plot(NA,
     ylab = expression(SoH^Down),
     xlab = sprintf("Weight matrix size"),
     xlim = c(7,43), ylim = c(0,1))
plotCI(x = mean$Group.1,
       y = mean$x,
       uiw = sd$x, err = "y", pch = 20, slty = 3, scol = "red", add = TRUE)
plotCI(x = meanG$Group.1,
       y = meanG$x,
       uiw = sdG$x, err = "y", pch = 20, slty = 3, scol = "blue", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
lines(meanG$Group.1, meanG$x, col = "blue", lw = 4)
legend("topleft", c("Focal G*", "G*") , fill = c("red", "blue"))

```

```{r aggregation, echo=FALSE, eval=TRUE, cache=TRUE, message=FALSE}
d3 <- read.csv("data/a_focal_trued3.csv")
d3global <- read.csv("data/a_focal_falsed3.csv")

mean <- aggregate(d3$Up, by = list(d3$Z + 1), mean)
sd <- aggregate(d3$Up, by = list(d3$Z + 1), sd)
meanG <- aggregate(d3global$Up, by = list(d3global$Z + 1), mean)
sdG <- aggregate(d3global$Up, by = list(d3global$Z + 1), sd)
plot(NA,
     ylab = expression(SoH^Up),
     xlab = sprintf("Aggregation levels"),
     xlim = c(2,6), ylim = c(0,1))
plotCI(x = mean$Group.1,
       y = mean$x,
       uiw = sd$x, err = "y", pch = 20, slty = 3, scol = "red", add = TRUE)
plotCI(x = meanG$Group.1,
       y = meanG$x,
       uiw = sdG$x, err = "y", pch = 20, slty = 3, scol = "blue", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
lines(meanG$Group.1, meanG$x, col = "blue", lw = 4)
legend("topleft", c("Focal G*","G*") , fill = c("red","blue"))

mean <- aggregate(d3$Down, by = list(d3$Z + 1), mean)
sd <- aggregate(d3$Down, by = list(d3$Z + 1), sd)
meanG <- aggregate(d3global$Down, by = list(d3global$Z + 1), mean)
sdG <- aggregate(d3global$Down, by = list(d3global$Z + 1), sd)
plot(NA, 
     ylab = expression(SoH^Down),
     xlab = sprintf("Aggregation levels"),
     xlim = c(2,6), ylim = c(0,1))
plotCI(x = mean$Group.1,
       y = mean$x,
       uiw = sd$x, err = "y", pch = 20, slty = 3, scol = "red", add = TRUE)
plotCI(x = meanG$Group.1,
       y = meanG$x,
       uiw = sdG$x, err = "y", pch = 20, slty = 3, scol = "blue", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
lines(meanG$Group.1, meanG$x, col = "blue", lw = 4)
legend("topleft", c("Focal G*","G*") , fill = c("red","blue"))
```

```{r focal, echo=FALSE, eval=TRUE, cache=TRUE, message=FALSE}

d3 <- read.csv("data/f_focal_trued3.csv")

mean <- aggregate(d3$Up, by = list(d3$F + 12), mean)
sd <- aggregate(d3$Up, by = list(d3$F + 12), sd)
plot(NA,
     ylab = expression(SoH^Up),
     xlab = "Focal matrix size",
     xlim = c(29, 137), ylim = c(0, 1))
plotCI(x = mean$Group.1, y = mean$x, uiw = sd$x, err = "y", pch = 20, slty = 3,
       scol = "red", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
legend("topleft", c("Focal G*") , fill = c("red"))

mean <- aggregate(d3$Down, by = list(d3$F + 12), mean)
sd <- aggregate(d3$Down, by = list(d3$F + 12), sd)
plot(NA,
     ylab = expression(SoH^Down),
     xlab = "Focal matrix size",
     xlim = c(29,137), ylim = c(0,1))
plotCI(x = mean$Group.1,
       y = mean$x,
       uiw = sd$x, err = "y", pch = 20, slty = 3, scol = "red", add = TRUE)
lines(mean$Group.1, mean$x, col = "red", lw = 4)
legend("topleft", c("Focal G*") , fill = c("red"))

```